\LARGE \bf Machine Translation for Tibetan Buddhist Texts

### 2021-04-15 paper/main.tex
Added more to the introduction

Our project aims to increase the accessibility of Tibetan religious texts by training a machine translation model specifically to translate religious Tibetan texts to English. We have trained our models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward. By creating these models, we hope to motivate scholars to study not only these texts but Tibetan as a whole. Currently, while there has been some interest in machine translation for Tibetan-Chinese and vice versa, the same interest has not been reflected in Tibetan-English machine translation. From a deep dive into the open-source Tibetan NLP material, we were unable to find any projects attempting to translate Tibetan-English nor could we find papers specifically discussing Tibetan-English machine translation. By having a readily available set of translated discourses could create an on-ramp for further language uses because increased access could lead to increased interest, especially in the English-speaking community.

>>>>

Our project aims to increase the accessibility of Tibetan religious texts by training a machine translation model specifically to translate religious Tibetan texts to English. We have trained our models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward. By creating these models, we hope to motivate scholars to study not only these texts but Tibetan as a whole. Having a readily available set of translated discourses could create an on-ramp for further language uses because increased access could lead to increased interest, especially in the English-speaking community. 
Currently, while there has been some interest in machine translation for Tibetan-Chinese and vice versa, the same interest has not been reflected in Tibetan-English machine translation. From a deep dive into the open-source Tibetan NLP material, we were unable to find any projects attempting to translate Tibetan-English nor could we find papers specifically discussing Tibetan-English machine translation. While our model tackles this difficult task in creating a translator for Tibetan-English, it is by no means a perfect or even accurate model. In the final iteration of our model, we were only able to achieve a 23.01% BLEU score, which means that our model has low-quality output. This is caused by several factors, from the difficulty to preprocess Tibetan sentences to the lack of Tibetan-English data freely available online and our overall knowledge of the Tibetan language. While it may seem that we were unable to fully accomplish our lofty goal of creating a Tibetan-English translator it’s important to note that we are treading in nearly uncharted territory. We hope that our endeavor may help future projects in this field and provide insight into the current state of Tibetan-English machine translation.
